---
title: "36-315 Lab 01"
author: "Jingyuan Xing"
date: "Due Saturday, August 31, 2019 (6:30pm)"
output:
  html_document: default
  pdf_document: default
---

## Lab 01: Hello, world! Introduction to R, RStudio, & RMarkdown

Google styleguide for R is used
***
***

# Problem 1

(15 points)

**Storing Your Work and Storing Your `R` Commands**: 

First, open the RStudio program on your computer. 

With R Markdown, you can simultaneously store your commands, execute
your commands, and generate an output file automatically. Because of
this, we will use R Markdown files (.Rmd) to store our code and
answers for all Lab and Homework assignments.
 
Use the Lab01.Rmd file as a template for your submission. Click "Knit
HTML" at the top (or Command+Shift+K on Mac). This should produce an
HTML page that executes the code you write. You can edit this .Rmd
file to produce your lab submission.

Notice how the text you write in the .Rmd file shows up in the output HTML file each time you click "Knit HTML". 

For more tips/tricks on how to format things in R Markdown, [see
here](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)


**Code Blocks**: 

When you open a new R Markdown file, you should see a block of code
(beginning with three [grave accent
marks](https://en.wikipedia.org/wiki/Grave_accent) -- similar to
apostrophes, but located next to the 1 key on a standard keyboard --
then an "`{r}`", then three more accent marks). This block should have
a light gray background color in R Studio. This is a code block! You
can type commands into this block, and they will be executed by `R`
and included in your output.

Let's test out the following command, the following command:

```{r}
# I am printing ‘Hello, World!’ in R Markdown
# My major is Statistics and Machine Learning
print("Hello, World! My name is Jingyuan Xing.")
```

Change the above code so that it outputs the following text: "Hello, World! My name is [your name]."


**Code Comments**: 

Comments should be used frequently when writing code to give insight
into what each piece of code is doing. To add a comment to your code,
start a new line with the `#` symbol.

Change the existing comment in your first code block so that it says
"I am printing 'Hello, World!' in R Markdown".

Add another comment that says what your major is. 

Notice that the text in the comments shows up in the code block in the
output file, but not in the actual output when you click "Knit
HTML". Comments exist to help you and others who read your code.


**Where is my output file?**

Find where you stored your [AndrewID]-315-Lab01.Rmd file on your computer. 

In the same directory, there should be a file called [AndrewID]-315-Lab01.html. 

Open it. It should automatically open in a browser, and it should contain the output. 


***
***


# Problem 2

(15 points)

**Getting Started**: Create a single code block that contains R commands to do all of the following:

a. $(6+3)*4 - 5$
b. $4^2$
c. $e^{-5}$ (hint: type ```help(exp)``` at the command line in R Studio
d. Let $y = 8$. Let $x = 5-6y$. Print $x$.
e. Repeat (d) for $y = 0$, $y = 1$, and $y = 5/6$.

```{r}
#a
(6+3)*4 - 5
#b
4^2
#c
exp(-5)
#d
findx <- function(y) {
  return(5-6*y)
  }
findx(8)
#e
findx(0)
findx(1)
findx(5/6)
```

***
***


# Problem 3

(15 points)

**Built-in Help and Documentation**: Using the `help()` and `help.search()` functions at the command line, use R to do the following:

a. Find the help documentation for the `quantile` function. This function takes a vector of numbers and computes quantiles for the vectors. What is the description of the `probs` argument?  

"probs" is: numeric vector of probabilities with values in [0,1]. (Values up to 2e-14 outside that range are accepted and moved to the nearby endpoint.)

b. Find the help documentation for the `mean` function. This function takes a vector of numbers and computes their average. What is the example code at the bottom of the help page?
  
The example code at the bottom of the help page:
```{r}
x <- c(0:10, 50)
xm <- mean(x)
c(xm, mean(x, trim = 0.10))
```
  
c. Use the help pages to find the name of the function in `R` that finds the standard deviation of a vector, and apply it to the data from b.
  
The standard deviation of data from b is: 
```{r}
partb <- c(xm, mean(x, trim = 0.10))
sd(partb)
```


***
***


# Problem 4

(15 points)

**Loading a library in R**: In `R`, there are many libraries or packages/groups of programs that are not permanently stored in `R`, so we have to load them when we want to use them. We can load the library into `R` by typing `library(library-name)` at the command line. (Sometimes we need to download the library first; more on this later.)

a. Load the `MASS` library into R. Open the help documentation for the MASS package. What is the official name of this `MASS` package? It may help to use `library(help = MASS)` to solve this problem.
  
```{r}
library(MASS)
```
The official name is "Modern Applied Statistics with S"
  
b. Load the `datasets` library. Find the help documentation for its `trees` dataset. Describe this dataset using information from the help pages.
  
```{r}
library(datasets)
```
The "trees" data set provides measurements of the diameter, height and volume of timber in 31 felled black cherry trees. Note that the diameter (in inches) is erroneously labelled Girth in the data. It is measured at 4 ft 6 in above the ground.

c. Load the `graphics` library into `R` and open its help documentation. This library is full of graphics/visualization functions that we may use in this class. Find a function that creates a 1-D Scatterplot. Describe its argument `x`.
  
```{r}
library(graphics)
```
The function that creates a 1-D scatterplot is "stripchart". It produces one dimensional scatter plots or dot plots of the given data. These plots are a good alternative to boxplots when sample sizes are small. Its argument x is the data from which the plots are to be produced. In the default method the data can be specified as a single numeric vector, or as list of numeric vectors, each corresponding to a component plot. In the formula method, a symbolic specification of the form y ~ g can be given, indicating the observations in the vector y are to be grouped according to the levels of the factor g. NAs are allowed in the data.

***
***


# Problem 5

(15 points)

**Installing Packages**

In this class, we'll use many `R` libraries ("packages") that are not currently installed on your computer or the lab computers. There are some security measures in place to prevent you from installing packages to CMU-wide repository of packages. Luckily, there are some ways around this.

a. At the command line / Console (NOT in a code block), type `install.packages("tidyverse")`. What happens? 

It worked.

If the package installed with no issues, you can skip the following parts! If you could not install the package, take the following steps:

b. Create a new directory on your computer called "36-315", and create a new sub-directory called "Packages". The filepath to this directory should be something like:

+ "C:/Users/YourName/Desktop/36-315/Packages" if you use Windows
+ "/Users/YourName/Desktop/36-315/Packages" if you use Mac
+ "..." if you are using the CMU cluster computers

c. In a code block, store the filepath in an object called `package_path`, e.g. `package_path <- "/Users/YourName/Desktop/36-315/Packages"`. Repeat this at the command line / Console as well.

d. In the same code block, include the following line of code: `.libPaths(c(package_path, .libPaths()))`

e. At the command line / Console (NOT in a code block), type `install.packages("tidyverse", lib = package_path)`. This should install the `tidyverse` package.

f. Finally, add `library(tidyverse)` to your code block from above. This should load the `tidyverse` package so that you can use it for the next problem.

Important note: Never install new packages in a code block. Always install new packages at the command line / Console. That is, the `install.packages()` function should never be in your submitted code. The `library()` function, however, should be in most of your submitted code.


***
***



# Problem 6

(25 points)

**Reading, Manipulating, and Plotting Data**

We will use the `read_csv()` function to 
read in a dataset from the internet. 
We'll use the 
dataset at [https://raw.githubusercontent.com/mateyneykov/315_code_data/master/data/bridges-pgh.csv] for this lab.



(Note that this data is in a GitHub repository called
"315_code_and_data". We'll use this repository for sharing various
datasets and code snippets throughout the semester.)



All code is provided for you for the following problems -- just uncomment the code in the code block below:

a. Load the `tidyverse` library.
b. Read in the "Pittsburgh Bridges" dataset.
c. Create a new column in the dataset that indicates whether the bridge crosses the Allegheny River ("yes" or "no").
d. Create a new column in the dataset that indicates whether the bridge is at least 3000 feet long ("long" or "short").
e. Create a well-labled bar plot of the `river` variable.

Although we provided the code for you here, it is important that you take some time to understand what the code does! Study it in detail, and be prepared to do similar tasks in subsequent assignments.

```{r, warning=FALSE, message=FALSE}
# Load the tidyverse library
library(tidyverse)

# Read in the data
bridges <- read_csv("https://raw.githubusercontent.com/mateyneykov/315_code_data/master/data/bridges-pgh.csv")



#Create the new variables
bridges <- mutate(bridges,
  over_allegheny = ifelse(river ==  "A", "yes", "no"),
  length_binary = ifelse(length >=  3000, "long", "short"))

#Create the bar plot
ggplot(bridges, aes(x = river)) +
geom_bar(fill = "darkblue") +
labs(
title = "Number of Bridges in Pittsburgh by River",
subtitle = "A = Allegheny River, M = Monongahela River, O = Ohio River",
x = "River",
y = "Number of Bridges",
caption = "Source: Pittsburgh Bridges Dataset"
)
```


***
***
***
***
***
***


