---
title: "36-315 Homework 07, Fall 2019"
author: "Jingyuan Xing"
date: "Due Wed Oct 30 2019 (11:00pm ET) on Canvas"
output: 
  html_document:
    toc:  true
    toc_float:  true
    code_folding:  show
---

#  Homework 07:  Higher Dimensional Continuous Data

***General instructions for all assignments***: 

+  Use this file as the template for your submission.  Delete the unnecessary text (e.g. this text, the problem statements, etc).  That said, keep the nicely formatted "Problem 1", "Problem 2", "a.", "b.", etc
+  Upload a single `R` Markdown file (named as:  [AndrewID]-315-HW07.Rmd -- e.g. "mneykov-315-HW07.Rmd") to the Homework 08 submission section on Canvas.  You do not need to upload the .html file.
+  The instructor and TAs will run your .Rmd file on their computers.  **If your .Rmd file does not knit on our computers,
you will get 0 points.**
+  Your file should contain the code to answer each question in its own code block.  Your code should produce plots/output that will be automatically embedded in the output (.html) file
+  Each answer must be supported by written statements (unless otherwise specified)
+  Include the name of anyone you collaborated with at the top of the assignment
+  Include the style guide you used below under Problem 0


***
***




#  Problem 0

**Organization, Themes, and HTML Output**

(5 points)

a.  For all problems in this assignment, organize your output as follows:

+  Organize each part of each problem into its own tab.  For Problems 2 and 3, the organization into tabs is already completed for you, giving you a template that you can use for the subsequent problems.
+  Use code folding for all code.  See [here](http://blog.rstudio.org/2016/03/21/rmarkdown-v0-9-5/) for how to do this.
+  Use a floating table of contents.
+  Suppress all warning messages in your output by using `warning = FALSE` and `message = FALSE` in every code block.


b.  For all problems in this assignment, adhere to the following guidelines for your `ggplot` theme and use of color:

+  Do not use the default `ggplot()` color scheme.
+  For any bar chart or histogram, outline the bars (e.g. with `color = "black"`).
+  Do not use both red and green in the same plot, since a large proportion of the population is red-green colorblind.
+  Try to only use three colors (at most) in your themes.  In previous assignments, many students are using different colors for the axes, axis ticks, axis labels, graph titles, grid lines, background, etc.  This is unnecessary and only serves to make your graphs more difficult to read.  Use a more concise color scheme.
+  Make sure you use a white or gray background (preferably light gray if you use gray).
+  Make sure that titles, labels, and axes are in dark colors, so that they contrast well with the light colored background.
+  Only use color when necessary and when it enhances your graph.  For example, if you have a univariate bar chart, there's no need to color the bars different colors, since this is redundant.
+  In general, try to keep your themes (and written answers) professional.  Remember, you should treat these assignments as professional reports.


c.  Treat your submission as a formal report:

+  Use complete sentences when answering questions.  
+  Answer in the context of the problem.  
+  Treat your submission more as a formal "report", where you are providing details analyses to answer the research questions asked in the problems.


Google Style Guide is used.


***
***


#  Problem 1

(3 points each)

**Text Annotations on Graphs**

Often, it can be nice to add text to graphs instead of more traditional geometries like points, bars, etc.  Let's learn how to do this in the context of scatterplots.

Read the help documentation for `geom_text()` [here](http://docs.ggplot2.org/current/geom_text.html) to help you solve the problems below.

This problem uses the `mpg` dataset which is loaded when you load `tidyverse`.

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
jingyuax_315_theme <-  theme_bw() + # White background, black and white theme
  theme(panel.background = element_rect(fill = "white",
                                colour = "lightgray",
                                size = 0.5, linetype = "solid"),
        text = element_text(size = 12, family = "Arial", face = "bold", color = "black"))
```

a.   Plot `cty` (y-axis) vs. `displ` (x-axis).  Use the `geom_text()` function instead of `geom_point()`, so that instead of points showing up in the scatterplot, text shows up instead.  Do this so that `manufacturer` of the car shows up as the text in the graph.

```{r, message=FALSE, warning=FALSE}
ggplot(mpg, aes(x = displ, y = cty))+
    geom_text(aes(label = manufacturer))+
    labs(title = "City Miles Per Gallon vs Engine Displacement",
         x = "Engine Displacement (L)",
         y = "City Miles Per Gallon")+
    jingyuax_315_theme
```

b.  Recreate the graph from (a), but this time, display the text at a 30 degree angle, and color the text according to the `drv` of the car  Don't forget to use a non-default color scheme!

```{r, message=FALSE, warning=FALSE}
ggplot(mpg, aes(x = displ, y = cty))+
    geom_text(aes(label = manufacturer, color = factor(drv)), angle = 30)+
    scale_colour_discrete(l = 40)+
    labs(title = "City Miles Per Gallon vs Engine Displacement",
         x = "Engine Displacement (L)",
         y = "City Miles Per Gallon",
         color = "Wheel Drive")+
    jingyuax_315_theme
```

c.  Recreate the graph from (b), but this time, size the text according to the `hwy`, and use a non-default font family. 

```{r, message=FALSE, warning=FALSE}
ggplot(mpg, aes(x = displ, y = cty))+
    geom_text(aes(label = manufacturer, color = factor(drv), size = hwy), angle = 30)+
    scale_colour_discrete(l = 40)+
    labs(title = "City Miles Per Gallon vs Engine Displacement",
         x = "Engine Displacement (L)",
         y = "City Miles Per Gallon",
         color = "Wheel Drive",
         size = "Highway Miles/Gallon")+
    jingyuax_315_theme
```

d.  Interpret your graph from (c).  How many variables are included in this graph?  What can you say about the relationships between any of the variables used in this graph?

Five variables are included in this graph, they are displ (engine displacement), cty (city miles per gallon), manufacturer, drv (front/back/four wheel drive), and hwy (highway miles per gallon). From the graph we can see that there is a moderate negative relationship between city miles per gallon and engine displacement, the relationship is not exactly linear, it looks more like a 1/x relationship. We can also see that highway miles per gallon is positively correlated with cty, because the larger y is, the larger the text size is; and highway miles per gallon is negatively correlated with engine displacement, because the smaller x is, the larger the text size is. front-wheel drive has the highest cty and lowest displ, 4-wheel drive is in the middle, and back-wheel drive has the lowest cty and highest displ.


***
***


#  Problem 2 {.tabset}

(2 points each)

**More Text Annotations on Graphs**

Here we will continue to work with the `mpg` dataset.

##  Part (a)

Create a bar chart of the `drv` variable.  Use the `geom_text()` function to add the counts in each bar **above** each bar in the bar chart.  E.g., `ggplot(...) + ... + geom_text(stat = "count", aes(y = ..count.., label = ..count..))`.  Adjust the `vjust` parameter in `geom_text()` in order to place the numbers in a more appropriate place.  Use a single non-default color for the bars.

```{r, warning = F, message = F}
ggplot(mpg, aes(x = drv))+
  geom_bar(fill = "light pink")+
  geom_text(stat = "count", aes(y = ..count.., label = ..count..), vjust = -0.5)+
  labs(title = "Bar Graph of Wheel Drive",
       x = "Wheel Drive")+
  jingyuax_315_theme
```

##  Part (b)

Create a bar chart of the `class` variable.  Use the `geom_text()` function to add the counts in each bar **in the middle** of each bar (i.e. halfway up each bar) in the bar chart.  Use a larger-than-default font size for the text.  Use a single non-default color for the bars, and change the color of the text so that it contrasts well with the color of the bars.  (Hint:  If `y = ..count..` works for making the text appear at the appropriate heights on the y-axis for part (a), how can you adjust this so that the text appears halfway up for each bar in this problem?)

```{r, warning = F, message = F}
ggplot(mpg, aes(x = class))+
  geom_bar(fill = "light pink")+
  geom_text(stat = "count", aes(y = 0.5 * (..count..), label = ..count..), vjust = 0.5, size = 5)+
  labs(title = "Bar Graph of Car Class",
       x = "Car Class")+
  jingyuax_315_theme
```

##  Part (c)

Repeat part (b), but use `label = scales::percent((..count..) / sum(..count..))` to put percentages on the plot instead of the raw count scale.  This is nice, since it allows you to quickly see both the percentages and the counts (height of the bars) in each category.

```{r, warning = F, message = F}
ggplot(mpg, aes(x = class))+
  geom_bar(fill = "light pink")+
  geom_text(stat = "count", aes(y = 0.5 * (..count..), label = scales::percent((..count..) / sum(..count..))), vjust = 0.5, size = 5)+
  labs(title = "Bar Graph of Car Class",
       x = "Car Class")+
  jingyuax_315_theme
```

***
***



#  Problem 3

**Hierarchical Clustering and Dendrograms**

*Load the `olive` dataset from Lab 07.*

```{r, message=F, warning=F}
olive <- read_csv("https://raw.githubusercontent.com/mateyneykov/315_code_data/master/data/olive_oil.csv")
```

There are several ways to create dendrograms in R.  Regardless of which dendrogram package you use, you'll first need to create the distance matrix corresponding to your dataset, and submit that distance matrix to hierarchical clustering.

a.  (1 point)  Create a subset of only the continuous variables (no `area` or `region`) from the `olive` dataset, called `olive_cont`.  Scale the dataset so that all of the variables are treated equally in the distance calculation (`olive_cont_scale`).  Now, create the distance matrix on this scaled version of the continuous variables (`dist_olive`).  

```{r, message=F, warning=F}
olive_cont <- olive[,-(1:2)]

normalized <- function(y) {
  x <- y[!is.na(y)]
  x <- (x - min(x)) / (max(x) - min(x))
  y[!is.na(y)] <- x
  return(y)
}
olive_cont_scale <- apply(olive_cont, 2, normalized)
dist_olive <- dist(olive_cont_scale)
```

b.  (2 points)  Submit your distance matrix to hierarchical clustering with the `hclust()` function.  Use complete linkage to determine when groups of observations should be linked:  `hclust(dist_olive, method = "complete")`.  Store this in an object called `hc_olive_complete`.  Type `names(hc_olive_complete)`.  What information is stored in the hierarchical clustering object?  What is `hc_olive_complete$method`?

```{r, message=FALSE, warning=FALSE}
hc_olive_complete <- hclust(dist_olive, method = "complete")
names(hc_olive_complete)
hc_olive_complete$method
```

The object hierarchical clustering is storing merge, height, order, labels, method, call, and dist.method. The hc_olive_complete$method is storing the method we are using when we build it, the method is complete.

c.  (2 points)  Plot your `hc_olive_complete` object using base R graphics:  `plot(hc_olive_complete)`.  One nice thing (or not-so-nice thing, depending on how you look at it) about base R graphics is that `plot()` will adapt to whatever kind of object it takes as input and show the corresponding plot in (what it deems to be) an appropriate way.  What kind of structure does `plot()` use to visualize hierarchical clustering results?

```{r, message=FALSE, warning=FALSE}
plot(hc_olive_complete)
```

plot() choose dendrograms to visualize hierarchical clustering results.

d.  (2 points) 
The resulting dendrogram specifies the distance at which groups of observations are linked in the hierarchical clustering results.  Groups linked at lower distances (lower heights on the dendrogram) are closer together / highly similar; groups linked at higher distances (higher heights on the dendrogram) are farther apart / highly dissimilar.  What, roughly, is the maximum distance at which two groups of observations are linked in this dataset?  The maximum distance at which two groups are linked represents the final link being completed in hierarchical clustering.  

e.  (2 points)  Dendrograms are tree-like structures.  Starting at the top of a dendrogram, if you follow the branches down to the "leaves" of the tree, you see a hierarchy of clustering results.  Each leaf in the tree represents a single observation in the dataset, so that there a total of $n$ leaves in the dendrogram (where $n$ is the number of observations in the dataset).  Roughly what proportion of the observations are in the two groups being linked at the final iteration of hierarchical clustering?  (An exact answer is not necessary here; just look at the graph and give a rough estimate of the number of leaves on the tree under each branch.)

Roughly 60% of the observations are in the left branch, and 40% of the observations are in the right branch of the final interation of hierarchical clustering.

f.  (2 points)  With hierarchical clustering, remember:  We actually have a _hierarchy_ of clustering results.  So, how do we actually get the clustering results?   There are two approaches with hierarchical clustering.  The first approach involves selecting the number of clusters that you want, and then "cutting" the tree an an appropriate height.  All observations linked below that height are considered to be in the same cluster.  All observations linked above that height are considered to be in different clusters.  To do this in R, use the `cutree()` function ("cut tree").  Return the cluster labels corresponding to the two-cluster solution by using `cutree(hc_olive_complete, k = 2)`, and store this in an object called `labels_complete_2`.  What type of object is `labels_complete_2`?  How many elements are contained in `labels_complete_2`?

```{r, message=FALSE, warning=FALSE}
labels_complete_2 <- cutree(hc_olive_complete, k = 2)
typeof(labels_complete_2)
length(labels_complete_2)
```

The type of labels_complete_2 is integer. There are 572 elements contained in labels_complete_2.

g.  (2 points)  Now, let's get the exact answer from part (e).  Type `table(labels_complete_2) / nrow(olive)`.  What proportion of the observations are in the two groups being linked at the final iteration of hierarchical clustering?

```{r, message=FALSE, warning=FALSE}
table(labels_complete_2) / nrow(olive)
```

The exact proportion of observations in the two groups being linked at the final iteration of hierarchical clustering is 37.41259% and 62.58741%, respectively.

h.  (8 points) Okay, now let's visualize how good the clustering
results are.  Create a new vector, called `labels_complete_3`, that
corresponds to the three-cluster solution.  Create the 2-dimensional
projection of `dist_olive` via multi-dimensional scaling.
Use the R command: cmdscale.
(Make sure you read the help file.)
Convert the
MDS results to a data.frame (`olive_mds`), and add the `area` and
`region` variables **as factors** to the `olive_mds` data.frame.
Now, add the `labels_complete_3`
variable (**as a factor**) to the `olive_mds` data.frame as well.
Finally, create a scatterplot with:

+  x = the first MDS coordinate
+  y = the second MDS coordinate
+  color = `area`
+  shape = `labels_complete_3`

That is, we're plotting the two MDS coordinates, coloring by the
`area` variable, and changing the type of points so that they are
integers corresponding to the labels from the three-cluster solution.
If you prefer to do this with `geom_text()` instead of the `shape`
parameter in `geom_point()`, please feel free to do so!  (It's
actually easier that way anyways!)

Do your cluster labels correspond to the `area` variable?  That is, is
it common for points with the same color (`area`) to also have the
point type (cluster label)?  Are there any particular `area`
categories that are split apart in the clustering results?

```{r, message=FALSE, warning=FALSE}
labels_complete_3 <- cutree(hc_olive_complete, k = 3)
olive_mds <- as.data.frame.matrix(cmdscale(dist_olive))
olive_mds <- olive_mds %>% mutate(area = as.factor(olive$area),
                                  region = as.factor(olive$region),
                                  labels_complete_3 = as.factor(labels_complete_3))
ggplot(olive_mds, aes(x = V1, y = V2))+
  geom_text(aes(color = area, label = labels_complete_3))+
  labs(title = "Scatterplot of The First vs The Second MDS coordinate",
       x = "The First MDS Coordinate",
       y = "The Second MDS Coordinate")+
  jingyuax_315_theme
```

The cluster labels doesn't exactly correspond to area. Points with the same color (area) don't also have the point type (cluster label). We can see that there are both cluster 1 and 2 with color red, and there are both cluster 1 and 3 with color blue. Cluster 1 is split into red and blue areas; cluster 2 is split into red and green areas.
